---
title: "Compile COPASI R Bindings From Source"
output:
  html_document:
    toc: yes
  html_notebook: default
---

# Initial checkout

Use this command to clone the repository:

```{bash, eval=FALSE}
git clone https://github.com/jpahle/CoRC
```

Alternatively via ssh:
```{bash, eval=FALSE}
git clone git@github.com:jpahle/CoRC.git
```

Once the repository is cloned, use:

```{bash, eval=FALSE}
git submodule update --init --recursive
```

This step makes sure the COPASI sources are cloned.

# macOS

## Dependencies for compilation of the COPASI sources

The following dependencies exist for compiling the sources:

xcode command line tools

```{bash, eval=FALSE}
xcode-select --install
```

Consider using homebrew for macOS to install other dependencies:

```{bash, eval=FALSE}
brew install cmake swig
```

## Full

Clears relevant folders before building.

```{bash, eval=FALSE}
cd development/copasi-source/
./build_darwin_clean.sh
cd ../../
```

## Quick

Requires copasi-dependencies to be compiled already.
Uses existing build but might cause issues when copasi_wrapper.cpp is old.

```{bash, eval=FALSE}
cd development/copasi-source/
./build_darwin_x64_quick.sh
cd ../../
```

## (Optionally) Build GUI

Instructions build on top of already compiled R bindings.

Cosider using homebrew for macOS intall qt5:

```{bash, eval=FALSE}
brew install qt5
```

Build dependencies:

```{bash, eval=FALSE}
cd development/copasi-source/copasi-dependencies/
export Qt5_Dir=/usr/local/opt/qt/lib/cmake/Qt5
BUILD_DIR=${PWD}/tmp_darwin_x64 \
	INSTALL_DIR=${PWD}/bin_darwin_x64 \
  ./createOSX-qt5.sh
cd ../../../
```

Build GUI:

```{bash, eval=FALSE}
cd development/copasi-source/build_copasi_r_binding/
export Qt5_Dir=/usr/local/opt/qt/lib/cmake/Qt5
cmake \
  -DBUILD_GUI=ON \
  -DSELECT_QT=Qt5 \
  -DQT5_USE_WEBENGINE=ON \
  .
make
cd ../../../
```

# Windows

On windows, get docker-ce.
On other systems, run a windows-docker-machine.
The windows-docker-machine uses vagrant and virtualbox to create a docker backend that can run windows containers.
Instructions here: https://github.com/StefanScherer/windows-docker-machine.
For evaluation only, consider the easy variant of this repo: 'vagrant up 2016-box'.

On macOS, preparation works like this:

```{bash, eval=FALSE}
export INSTALL_DIR = "~/"
brew cask install docker-ce vagrant virtualbox
cd $INSTALL_DIR
git clone https://github.com/StefanScherer/windows-docker-machine
cd windows-docker-machine
vagrant up 2016-box
```

You now run a headless VM in the background which can host the Docker container.

Confirm that the '2016-box' machine is listed:

```{bash, eval=FALSE}
docker-machine ls
```

Use this to stop from install folder when done:

```{bash, eval=FALSE}
vagrant halt 2016-box
vagrant destroy 2016-box
```

## Full

Clears relevant folders before building.

```{bash, eval=FALSE}
cd development/copasi-source/
./d_build_windows_clean.sh
cd ../../
```

alternatives:
  on windows: d_build_windows_clean.ps1
  without docker: build_windows_clean.bat (see windows Dockerfile for system requirements)

## Quick

Requires copasi-dependencies to be compiled already.

```{bash, eval=FALSE}
cd development/copasi-source/
./d_build_windows_x64_quick.sh
cd ../../
```

alternatives:
  on windows: d_build_windows_quick.ps1
  without docker: build_windows_quick.bat (see windows Dockerfile for system requirements)

# Linux

For linux we generally use docker.
Install docker CE and get docker permissions for your user.

## Full

Clears relevant folders before building.

```{bash, eval=FALSE}
cd development/copasi-source/
./d_build_linux_clean.sh
cd ../../
```

alternatives:
  on windows: d_build_linux_clean.ps1
  without docker: build_linux_clean.sh (see linux Dockerfile for system requirements)

## Quick

Requires copasi-dependencies to be compiled already.

```{bash, eval=FALSE}
cd development/copasi-source/
./d_build_linux_x64_quick.sh
cd ../../
```

alternatives:
  on windows: d_build_linux_quick.ps1
  without docker: build_linux_quick.sh (see linux Dockerfile for system requirements)

# Copying the relevant file into the package repository

Will also patch the swig wrapper (COPASI.R) as per patch-swig-wrapper.R

```{r, eval=FALSE}
source("../development/update-swig-in-package.R", chdir = TRUE)
```

# Installing the package

```{r, eval=FALSE}
devtools::install("..")
```

# Getting the current hash for the binary (update it in CoRC.R)

```{r, eval=FALSE}
for (file in list.files("copasi-source/libs", full.names = T))
  cat(file, "\t", digest::digest(file, algo = "sha256", file = TRUE), "\n")
```
