---
title: "Compile Copasi R Bindings From Source"
output:
  html_document:
    toc: yes
  html_notebook: default
---

# Initial checkout

Use this command to clone the repository:

```{bash, eval=FALSE}
git clone https://github.com/jpahle/CoRC
```

Alternatively via ssh:
```{bash, eval=FALSE}
git clone git@github.com:jpahle/CoRC.git
```

Once the repository is cloned, use:

```{bash, eval=FALSE}
git submodule update --init --recursive
```

This step makes sure the Copasi sources are cloned.

# Dependencies for compilation of the copasi sources

## macOS

The following dependencies exist for compiling the sources:

xcode command line tools

```{bash, eval=FALSE}
xcode-select --install
```

Cosider using homebrew for MacOS to install other dependencies:

cmake, swig

```{bash, eval=FALSE}
brew install cmake swig
```

# Compiling copasi dependencies

## macOS

```{bash, eval=FALSE}
cd ../development/copasi-source/copasi-dependencies/
./createOSX.sh
cd ../../../documentation/
```

# Compiling copasi from source

```{bash, eval=FALSE}
cd ../development/copasi-source/
mkdir build_copasi_r_bindings
cd build_copasi_r_bindings/
cmake -DBUILD_GUI=OFF -DBUILD_SE=OFF -DENABLE_R=ON -DCOPASI_DEPENDENCY_DIR=../copasi-dependencies/bin ../COPASI
make
cd ../../../documentation/
```

# Copying the relevant file into the package repository

```{r}
source("../development/update-swig-in-package.R", chdir = TRUE)
```

# Installing the package

```{r}
devtools::install("..")
```

# Getting the current hash for the binary (update it in CoRC.R)

```{r}
digest::digest("copasi-source/build_copasi_r_bindings/copasi/bindings/R/COPASI.so", algo = "sha256", file = TRUE)
```

