---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(1)
```

# <img src="man/figures/logo.svg" alt="CoRC logo" width="200"> --- the <b>Co</b>pasi <b>R</b> <b>C</b>onnector

## Overview

CoRC, the Copasi R Connector, links the Complex Pathway Simulator COPASI ([copasi.org](http://copasi.org)) and the (statistical) programming environment R ([r-project.org](http://r-project.org)). It provides easy access to the powerful biochemical model editing, simulation and analysis backend of Copasi from the convenient R command line interface. This allows the user to develop elaborate scripts and workflows for analyses that would require a great deal of tedious manual work otherwise. These scripts can then be run interactively from the R command line interface or send to cluster or cloud facilities for more demanding calculations.

CoRC features:

* high-level API for Copasi in the R language
* ...

You can get more information or download CoRC from <https://github.com/jpahle/CoRC/>.

CoRC comes with the Artistic License 2.0. By using CoRC you agree to this license.

## Installation

Install the development version from GitHub.

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jpahle/CoRC")
```

## Usage

```{r example}
library(CoRC)
loadExamples(1)
runTimeCourse()
```
