---
title: "Compile COPASI R Bindings From Source"
output:
  html_document:
    toc: yes
  html_notebook: default
---

# R Version specificity

For now, it seems that all version of R 3.x have been somewhat compatible.
There has been cases though, where for example R 3.5 can't run binaries compiled against R 3.4.
So to ensure maximum compatiblity, you should ensure to always compile against the latest version of R.
Consider using snapshots for new releases after code freeze.
For mac, install manually. For win and linux, adjust the Dockerfile.

# Initial checkout

Use this command to clone the repository:

```{bash, eval=FALSE}
git clone https://github.com/jpahle/CoRC
```

Alternatively via ssh:
```{bash, eval=FALSE}
git clone git@github.com:jpahle/CoRC.git
```

Once the repository is cloned, use:

```{bash, eval=FALSE}
git submodule update --init --recursive
```

This step makes sure the COPASI sources are cloned.

# macOS

## Dependencies for compilation of the COPASI sources

The following dependencies exist for compiling the sources:

xcode command line tools

```{bash, eval=FALSE}
xcode-select --install
```

Consider using homebrew for macOS to install other dependencies:

```{bash, eval=FALSE}
brew install cmake swig git
```

## Full

Clears relevant folders before building.

```{bash, eval=FALSE}
cd copasi-source/
./build_darwin_x64_clean.sh
cd ../
```

## Quick

Requires copasi-dependencies to be compiled already.
Uses existing build but might cause issues when copasi_wrapper.cpp is old.

```{bash, eval=FALSE}
cd copasi-source/
./build_darwin_x64_quick.sh
cd ../
```

## (Optionally) Build GUI

Instructions build on top of already compiled R bindings.

Cosider using homebrew for macOS intall qt5:

```{bash, eval=FALSE}
brew install qt
```

Build dependencies:

```{bash, eval=FALSE}
cd copasi-source/copasi-dependencies/
export PATH="/usr/local/opt/qt/bin:$PATH"
BUILD_DIR=${PWD}/tmp_darwin_x64 \
	INSTALL_DIR=${PWD}/bin_darwin_x64 \
  ./createOSX-qt5.sh
cd ../../
```

Build GUI:

```{bash, eval=FALSE}
cd copasi-source/corc_darwin_x64/
export PATH="/usr/local/opt/qt/bin:$PATH"
cmake \
  -DBUILD_GUI=ON \
  -DSELECT_QT=Qt5 \
  -DQT5_USE_WEBENGINE=ON \
  .
make
cd ../../
```

# Windows

On windows, get docker-ce.
On other systems, run a windows-docker-machine.
The windows-docker-machine uses vagrant and virtualbox to create a docker backend that can run windows containers.
Instructions here: https://github.com/StefanScherer/windows-docker-machine.
For evaluation only, consider the easy variant of this repo: 'vagrant up 2016-box'.

On macOS, preparation works like this:

```{bash, eval=FALSE}
export INSTALL_DIR = "~/"
brew cask install docker-ce vagrant virtualbox
cd $INSTALL_DIR
git clone https://github.com/StefanScherer/windows-docker-machine
cd windows-docker-machine
vagrant up 2016-box
```

You now run a headless VM in the background which can host the Docker container.

Confirm that the '2016-box' machine is listed:

```{bash, eval=FALSE}
docker-machine ls
```

Use this to stop from install folder when done:

```{bash, eval=FALSE}
vagrant halt 2016-box
vagrant destroy 2016-box
```

## Full

Clears relevant folders before building.

```{bash, eval=FALSE}
cd copasi-source/
./d_build_windows_x64_clean.sh
cd ../
```

alternatives:
  on windows: d_build_windows_x64_clean.ps1
  without docker: build_windows_x64_clean.bat (see windows Dockerfile for system requirements)

## Quick

Requires copasi-dependencies to be compiled already.

```{bash, eval=FALSE}
cd copasi-source/
./d_build_windows_x64_quick.sh
cd ../
```

alternatives:
  on windows: d_build_windows_x64_quick.ps1
  without docker: build_windows_x64_quick.bat (see windows Dockerfile for system requirements)

# Linux

For linux we generally use docker.
Install docker CE and get docker permissions for your user.
The manylinux2010 container is the basis for the linux binaries.

## Full

Clears relevant folders before building.

```{bash, eval=FALSE}
cd copasi-source/
./d_build_manylinux2010_x64_clean.sh
cd ../
```

alternatives:
  on windows: d_build_manylinux2010_x64_clean.ps1
  without docker: build_linux_clean.sh (see linux Dockerfile for system requirements)

## Quick

Requires copasi-dependencies to be compiled already.

```{bash, eval=FALSE}
cd copasi-source/
./d_build_manylinux2010_x64_quick.sh
cd ../
```

alternatives:
  on windows: d_build_manylinux2010_x64_quick.ps1
  without docker: build_linux_quick.sh (see linux Dockerfile for system requirements)

# Copying the relevant file into the package repository

Will also patch the swig wrapper (COPASI.R) as per patch-swig-wrapper.R

```{r, eval=FALSE}
source("../dev/update-swig-in-package.R", local = new.env(), chdir = TRUE)
```

# Installing the package

Ensures, that the installed package has a working lib

```{r, eval=FALSE}
devtools::install("..")
CoRC::getCopasi(file.path("..", "libs", paste0("COPASI", .Platform$dynlib.ext)))
```

# Getting the current hash for the binary (update it in CoRC.R)

```{r, eval=FALSE}
for (file in list.files("copasi-source/libs", full.names = T))
  cat(file, "\t", digest::digest(file, algo = "sha256", file = TRUE), "\n")
```
