---
title: "CoRC - the Copasi R Connector"
author: "Jonas Förster & Jürgen Pahle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# <img src="logo/CoRC.svg" alt="CoRC logo" width="200"> --- the <b>Co</b>pasi <b>R</b> <b>C</b>onnector

CoRC, the Copasi R Connector, links the Complex Pathway Simulator COPASI ([copasi.org](http://copasi.org)) and the (statistical) programming environment R ([r-project.org](http://r-project.org)). It provides easy access to the powerful biochemical model editing, simulation and analysis backend of Copasi from the convenient R command line interface. This allows the user to develop elaborate scripts and workflows for analyses that would require a great deal of tedious manual work otherwise. These scripts can then be run interactively from the R command line interface or send to cluster or cloud facilities for more demanding calculations.

CoRC features:

* high-level API for Copasi in the R language
* ...

You can get more information or download CoRC from <https://github.com/jpahle/CoRC/>.


# Installation

You can install CoRC from [github](http://github.com) with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jpahle/CoRC")
CoRC::getCopasi()
```

CoRC comes with the Artistic License 2.0. By using CoRC you agree to this license.


# Examples, case-studies and workflows

to be extended...

```{r}
require(CoRC)
require(ggplot2)
loadSBML("http://www.ebi.ac.uk/biomodels-main/download?mid=BIOMD0000000329")
timecourse <- runTimeCourse(50, 0.1)
autoplot(timecourse)
```
