---
title: "Model Entity Management"
author: "Jonas Förster & Jürgen Pahle"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model Entity Management}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.asp = 10/16,
  out.width = "100%",
  error = TRUE,
  warning = FALSE,
  message = FALSE
)
set.seed(1)
```

# Initial Setup

```{r setup}
library(CoRC)
loadExamples(2)
species()
```

# Identifying entities

There are various types of model entities supported for identification by a given string.
The strings are matched against the available names of e.g. species, compartments, or global quantities.
The fuctions are listed below and return entitiy `key`s when entities are identified.

```{r}
species("CheA")
compartment("comp")
quantity("a")
reaction("Che B phosph")
parameter("4).kR")
kinfunction("Michaelis-Menten")
event()
```

There are `_strict` variants that will always return a vector of `key`s of the same length as the input or throw an error.
Accordingly, the behavior for `NA` or non-uniquely matching strings is to throw an error.
This allows the `_strict` variants to uniquely identify species by giving the full name, when the other methods might then still return multiple matches.

```{r}
species("CheY")
species_strict("CheY")
species_strict("CheYp")
species_strict(c("CheY", "CheYp"))
species_strict("Che")
species_strict(c(NA, "CheY"))
species_strict("failure")
```

CoRC makes use of the regex functionality of the 'stringr' R package.
You can therefore wrap any string via `regex(string)` to enable regex matching.

```{r}
species(regex("Che.p"))
```

In cases, where a function argument `key` is clear to always expect the same kind of entity, identification can also be handled by the function itself.
These methods will always use the `_strict` variants to reliably match species.

```{r}
getSpecies(key = species("CheY"))$name
getSpecies(key = species_strict("CheY"))$name
getSpecies(key = "CheY")$name
getSpecies("CheY")$name
getSpecies("Che")$name
getSpecies(species(regex("Che.p")))$name
```

# Using References

The identifying functions can also return value references if specified.
Such references identify specific values, of which most COPASI entities can have multiple.
This allows to differentiate for example between 'Concentration' and 'Initial Concentration' of a species.
References are needed e.g. for definition of optimization task parameters.
They also allow for building expressions, which are needed e.g. in value assignments and the optimization task.
Expressions in CoRC behave similar to how they work in COPASI.

```{r}
species("CheY", reference = "InitialConcentration")
```

These functions however can not give any clues about what references are available.

```{r}
species("CheY", reference = "failure")
```

To help with this, functions implemented to get model values have similar functions that return references instead of values.

```{r}
str(getSpecies("CheY"))
str(getSpeciesReferences("CheY"))
```

# Building expressions

Any expression can easily be built as a string.
In the following example, several related species are selected via a REGEX rule and the reference to their initial concentration is requested.
The expression is then built by concatenating these references with a plus sign.
`getValue` calculates this expressions value and thereby the sum of the individual species.

```{r}
matching_init_conc <- species(regex("^T[0-4]"), reference = "InitialConcentration")
matching_init_conc
expression <- paste0(matching_init_conc, collapse = " + ")
expression
getValue(expression)
```

# Using numbers for referencing

Numbers can in principle be used by getting all entities and picking an index from that vector:

```{r}
species()[5]
getSpecies(species()[5])$name
```

It is under consideration to allow for a simpler method:

```{r}
species(5)
species_strict(5)
getSpecies(5)$name
```

# Entity modification

For most kinds of entities there is a finder function (e.g. `species`), a getter function (e.g. `getSpecies`) and a setter function (e.g. `setSpecies`).

Another way to access model values is by evaluating expressions (e.g. `getValue`):

```{r}
getSpecies("CheYp")$initial_concentration

setSpecies("CheYp", initial_concentration = 1)
getSpecies("CheYp")$initial_concentration

yp_init_expression <- species("CheYp", reference = "InitialConcentration")
yp_init_expression
getValue(yp_init_expression)
```
