---
title: "Managing Model Entities"
author: "Jonas Förster & Jürgen Pahle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing Model Entities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  out.width = "100%",
  error = TRUE,
  warning = FALSE,
  message = FALSE
)
set.seed(1)
```

# Initial Setup

```{r setup}
library(CoRC)
loadExamples(2)
species()
```

# Identifying entities

There are at the moment six different types of model entities supported for identification by string.
The strings are matched against the available entity names.
These functions return an entity `key` when a species is identified.

```{r}
species("CheA")
compartment("comp")
quantity("a")
reaction("Che B phosph")
parameter("4).kR")
kinfunction("Michaelis-Menten")
```

There are `_strict` variants that will always return a vector of `key`s of the same length as the input or throw an error.
Accordingly, the behavior for `NA` or non-uniquely matching strings is to throw an error.
This allows the `_strict` variants to uniquely identify species by giving the full name, when the other methods might then still return multiple matches.

```{r}
species("CheY")
species_strict("CheY")
species_strict("CheYp")
species_strict(c("CheY", "CheYp"))
species_strict("Che")
species_strict(c(NA, "CheY"))
species_strict("failure")
```

CoRC makes use of the regex functionality of the stringr package.
You can therefore wrap any string via `regex(string)` to enable regex matching.

```{r}
species(regex("Che.p"))
```

In cases, where a function argument `key` is clear to always expect the same kind of entity, identification can also be handled by the function itself.
These methods will always use the `_strict` variants to reliably match species.

```{r}
getSpecies(key = species("CheY"))$name
getSpecies(key = species_strict("CheY"))$name
getSpecies(key = "CheY")$name
getSpecies("CheY")$name
getSpecies("Che")$name
getSpecies(species(regex("Che.p")))$name
```

# Using References

The identifying functions can also return value references if specified.
Such references identify specific values, of which most copasi entities can have multiple.
This allows to differentiate for example between 'Concentration' and 'Initial Concentration' of a species.
References are needed e.g. for definition of optimization task parameters.
They also allow for building expressions, which are needed e.g. in value assignments and the optimization task.
Expressions in CoRC behave similar to how they work in Copasi.

```{r}
species("CheY", reference = "InitialConcentration")
```

These functions however can not give any clues about what references are available.

```{r}
species("CheY", reference = "failure")
```

To help with this, functions implemented to get model values have similar functions that return references instead of values.

```{r}
str(getSpecies("CheY"))
str(getSpeciesReferences("CheY"))
```

# Building expressions

Any expression can easily be built as a string.
`getValue` calculates an expression value.

```{r}
expression <- paste0(species(regex("^T[0-4]"), reference = "InitialConcentration"), collapse = " + ")
expression
getValue(expression)
```

# Extension suggestion

Using numbers for referencing:

```{r}
species()[5]
getSpecies(species()[5])$name
```

Maybe I should allow integers to be passed as keys:

```{r}
species(5)
species_strict(5)
getSpecies(5)$name
```

# Entity modification

For most kinds of entities there is a finder function (e.g. `species`), a getter function (e.g. `getSpecies`) and a setter function (e.g. `setSpecies`).

Other ways to access model values is by evaluating expressions (e.g. `getValue`).

```{r}
getSpecies("CheYp")$initial_concentration

setSpecies("CheYp", initial_concentration = 1)
getSpecies("CheYp")$initial_concentration

yp_init_expression <- species("CheYp", reference = "InitialConcentration")
yp_init_expression
getValue(yp_init_expression)
```
