---
title: "Case-Studies and Workflows"
author: "Jonas Förster & Jürgen Pahle"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.asp = 3/4,
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
set.seed(1)
```

This article contains case-studies illustrating the benefits of implementing workflows with CoRC.
Example code to get started can be found in the tutorial articles on [model entitiy management](https://jpahle.github.io/CoRC/articles/entity_management.html), [task management](https://jpahle.github.io/CoRC/articles/task_management.html) and [model building](https://jpahle.github.io/CoRC/articles/model_building.html).

# Initial Setup

```{r setup, results="hide"}
library(tidyverse)
library(parallel)
library(CoRC)

# helper to run tasks in parallel on all cores
mapInParallel <- function(data, fun, ..., .prep = {}) {
  cl <- makeCluster(detectCores())
  clusterCall(cl = cl, fun = eval, .prep, env = .GlobalEnv)
  result <- parLapplyLB(cl = cl, X = data, fun = as_mapper(fun), ..., chunk.size = 50)
  stopCluster(cl)
  result
}
```

# 3D Trajectory Plot of a Calcium Model

This example loads the [Kummer2000 - Oscillations in Calcium Signalling](https://www.ebi.ac.uk/biomodels-main/Biomd0000000329) model.
The model has 3 species which oscillate.
These oscialltions can be visualized as a trajectory through a 3D space.
The example does this once in a deterministic and once in a stochatic fashion.

```{r}
loadSBML("http://www.ebi.ac.uk/biomodels-main/download?mid=BIOMD0000000329")

# Run 24 sec (2 Periods)
setTimeCourseSettings(24, intervals = 10000)

timeseries <- list(
  deterministic = runTimeCourse()$result,
  stochastic = runTimeCourse(method = list(method = "directMethod", use_random_seed = T, random_seed = 1))$result
)

# Create the same plot for both timeseries
plots <- map(
  timeseries,
  plotly::plot_ly,
  type = "scatter3d",
  mode = "lines",
  x = ~ `G-alpha`,
  y = ~ activePLC,
  z = ~ Calcium,
  color = ~ Time
)

unloadModel()

plots$deterministic
plots$stochastic
```

# Statistics of Repeated Stochastic Simulations

This implements an example from the [Condor-COPASI paper](https://doi.org/10.1186/1752-0509-6-91).
The example illustrates advantages of parallel processing.

```{r}
# Run 1000 stochastic time series possibly in parallel
loadModel("https://raw.githubusercontent.com/copasi/condor-copasi/master/examples/stochastic_test.cps")

# timeseries <- 1:1000 %>% map(~ runTimeCourse()$result)
timeseries <-
  # Defines parallel evaluation:
  mapInParallel(
    # perpare worker (.prep),
    .prep = quote({
      library(CoRC)
      loadModel("https://raw.githubusercontent.com/copasi/condor-copasi/master/examples/stochastic_test.cps")
      setTimeCourseSettings(method = list(method = "directMethod", use_random_seed = T))
    }),
    # iteration data (1000 random seeds) is given as .x to iteration formula,
    1:1000,
    # iteration code (formula ~ {.x})
    ~ runTimeCourse(method = list(random_seed = .x))$result
  )

# Combine all results and reshape the data
plotdata <-
  timeseries %>%
  bind_rows() %>%
  group_by(Time) %>%
  # calculate mean and sd for all time points
  summarise_all(list(mean = mean, sd = sd)) %>%
  # gather all values so the column "name" identifies "a_mean", "b_sd" etc.
  pivot_longer(-Time) %>%
  # split up information on species (a,b,c) and type of value (mean, sd)
  separate(name, c("species", "type"), "_") %>%
  pivot_wider(names_from = type, values_from = value)

print(plotdata, n = 6)

plot <-
  ggplot(data = plotdata, aes(x = Time, y = mean, group = species, tt_sd = sd)) +
  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd, fill = species), alpha = 1 / 4) +
  geom_line(aes(color = species)) +
  guides(fill = "none") +
  labs(
    x = paste0("Time (", getTimeUnit(), ")"),
    y = paste0("Concentration (", getQuantityUnit(), ")"),
    color = "Species"
  )

unloadModel()

plotly::ggplotly(plot, tooltip = c("group", "x", "y", "tt_sd"))
```

# Parameter Scan

## 2D Scan Over the Cartesian Product of Two Species Concentration Vectors

This implements an example from the [Mendes2009 paper](https://doi.org/10.1007/978-1-59745-525-1_2) on COPASI use cases.

```{r}
loadSBML("https://www.ebi.ac.uk/biomodels-main/download?mid=BIOMD0000000068")

setSteadyStateSettings(calculate_jacobian = FALSE, perform_stability_analysis = FALSE)

# Cartesian product of the input values
scan <- cross_df(
  list(
    cysteine = 0.3 * 10 ^ seq(0, 3, length.out = 6),
    adomed = seq(0, 100, length.out = 51)
  )
)

scan <-
  scan %>%
  rowwise() %>%
  mutate(
    # Calculate steady state fluxes for every row
    ss_fluxes = {
      setSpecies("Cysteine", initial_concentration = cysteine)
      setSpecies("adenosyl", initial_concentration = adomed)
      ss <- runSteadyState()
      stopifnot(ss$result == "found")
      list(ss$reactions$flux)
    },
    # The second flux is CGS
    CGS = ss_fluxes[2],
    # The third flux is TS
    TS = ss_fluxes[3]
  )

plot <-
  ggplot(data = scan, aes(x = adomed, group = cysteine)) +
  geom_point(aes(y = CGS, color = "CGS")) +
  geom_point(aes(y = TS, color = "TS")) +
  geom_line(aes(y = CGS, color = "CGS")) +
  geom_line(aes(y = TS, color = "TS")) +
  labs(
    x = paste0("Adomed (", getQuantityUnit(), ")"),
    y = paste0("Flux (", getQuantityUnit(), " / ", getTimeUnit(), ")"),
    color = "Species"
  )

unloadModel()

plotly::ggplotly(plot)
```

## 2D Scan over Random Concentrations of Two Species

This implements an example from the [Mendes2009 paper](https://doi.org/10.1007/978-1-59745-525-1_2) on COPASI use cases.
It is in many ways similar to the previous example but is written to run parallelized.

```{r}
loadSBML("https://www.ebi.ac.uk/biomodels-main/download?mid=BIOMD0000000068")

# 10000 repeats of steady state task with random cysteine and adomed
scan <-
  # Defines parallel evaluation:
  mapInParallel(
    # perpare worker (.prep),
    .prep = quote({
      library(CoRC)
      loadSBML("https://www.ebi.ac.uk/biomodels-main/download?mid=BIOMD0000000068")
      setSteadyStateSettings(calculate_jacobian = FALSE, perform_stability_analysis = FALSE)
    }),
    # iteration data (10000 random seeds) is given as .x to iteration formula,
    1:10000,
    # iteration code (formula ~ {.x})
    ~ {
      set.seed(.x)
      cysteine <- 0.3 * 10 ^ runif(1L, min = 0, max = 3)
      adomed <- runif(1L, min = 0, max = 100)
      setSpecies(
        key = c("Cysteine", "adenosyl"),
        initial_concentration = c(cysteine, adomed)
      )
      ss <- runSteadyState()
      stopifnot(ss$result == "found")
      list(
        cysteine = cysteine,
        adomed = adomed,
        CGS = ss$reactions$flux[2],
        TS = ss$reactions$flux[3]
      )
    }
  )

# Combine all results and reshape the data
plotdata <-
  scan %>%
  bind_rows() %>%
  pivot_longer(c(CGS, TS), names_to = "reaction", values_to = "flux")

plot <-
  ggplot(data = plotdata, aes(x = adomed, y = flux, group = reaction, tt_cys = cysteine)) +
  geom_point(aes(color = reaction), alpha = 1 / 10, size = 3 / 4) +
  labs(
    x = paste0("Adomed (", getQuantityUnit(), ")"),
    y = paste0("Flux (", getQuantityUnit(), " / ", getTimeUnit(), ")"),
    color = "Species"
  )

unloadModel()

plotly::ggplotly(plot, tooltip = c("tt_cys", "x", "y"))
```

# Parameter Estimation

This implements an example from the [Mendes2009 paper](https://doi.org/10.1007/978-1-59745-525-1_2) on COPASI use cases.

```{r}
loadSBML("http://www.ebi.ac.uk/biomodels-main/download?mid=BIOMD0000000010")

# get timeseries for the record
data_before <- runTimeCourse(1000, 1)$result

# read experimental data
data_experimental <-
  read_tsv("data/MAPKdata.txt") %>%
  rename(Time = time, `Mos-P` = "MAPKKK-P", `Erk2-P` = "MAPK-P")

# define the experiments for COPASI
fit_experiments <- defineExperiments(
  data = data_experimental,
  type = c("time", "dependent", "dependent"),
  mapping = c(NA, "{[Mos-P]}", "{[Erk2-P]}"),
  weight_method = "mean_square"
)

# define the parameters for COPASI
fit_parameters <-
  map(
    parameter_strict(regex(c("V1$", "V2$", "V5$", "V6$", "V9$", "V10$"))),
    ~ {
      val <- getParameters(.x)$value
      defineParameterEstimationParameter(parameter(.x, "Value"), start_value = val, lower_bound = val * 0.1, upper_bound = val * 1.9)
    }
  )

result <-
  runParameterEstimation(
    parameters = fit_parameters,
    experiments = fit_experiments,
    method = list(
      method = "LevenbergMarquardt",
      log_verbosity = 2
    ),
    update_model = TRUE
  )

# get timeseries for the record
data_after <- runTimeCourse(1000, 1)$result

plots <- list(
  `Erk2-P` =
    ggplot(mapping = aes(x = Time, y = `Erk2-P`)) +
    geom_point(data = data_experimental, aes(color = "experimental")) +
    geom_line(data = data_before, aes(color = "before")) +
    geom_line(data = data_after, aes(color = "after")) +
    scale_color_manual(values = c(before = "red", after = "green", experimental = "black")) +
    labs(
      x = paste0("Time (", getTimeUnit(), ")"),
      y = paste0("Erk2-P (", getQuantityUnit(), ")"),
      color = "Series"
    ),
  `Mos-P` =
    ggplot(mapping = aes(x = Time, y = `Mos-P`)) +
    geom_point(data = data_experimental, aes(color = "experimental")) +
    geom_line(data = data_before, aes(color = "before")) +
    geom_line(data = data_after, aes(color = "after")) +
    scale_color_manual(values = c(before = "red", after = "green", experimental = "black")) +
    labs(
      x = paste0("Time (", getTimeUnit(), ")"),
      y = paste0("Mos-P (", getQuantityUnit(), ")"),
      color = "Series"
    )
)

unloadModel()

result$fitted_values
result$parameters
result$protocol

plotly::ggplotly(plots$`Erk2-P`)
plotly::ggplotly(plots$`Mos-P`)
```
