---
title: "Handling Model Entities"
author: "Jonas Förster & Jürgen Pahle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling Model Entities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  out.width = "100%",
  error = TRUE,
  warning = FALSE,
  message = FALSE
)
set.seed(1)
```

## Initial Setup

```{r setup}
library(tidyverse)
library(CoRC)
loadExamples(2)
getSpecies()
```

## Identifying entities

There are at the moment 5 different types of model entities supported for identification by string.
The strings are matched against the available ObjectDisplayNames.

```{r}
species("CheA")
compartment("comp")
quantity("a")
reaction("Che B phosph")
parameter("4).kR")
```

These functions return ObjectDisplayNames when a species is identified.

There are vectorized variants that will always return a vector of identified entities of the same length or throw an error.
This means the behavior for 'NA' or non-uniquely matching strings is to throw an error.
This allows to uniquely identify species by giving the full name, when the other methods might then still return multiple matches.

```{r}
species("CheY")
species_strict("CheY")
species_strict("CheYp")
species_strict(c("CheY", "CheYp"))
species_strict("Che")
species_strict(c(NA, "CheY"))
species_strict("failure")
```

CoRC makes use of the regex functionality of the stringr package.
You can therefore wrap any string via \code{regex(string)} to enable regex matching.

```{r}
species(regex("Che.p"))
```

In cases, where a key is clear to always expect the same kind of entity, identification can also be handled by the function itself.
These methods will always use the strict variants to allow reliably matching species.

```{r}
getSpecies(key = species("CheY"))
getSpecies(key = species_strict("CheY"))
getSpecies(key = "CheY")
getSpecies("CheY")
getSpecies("Che")
getSpecies(species(regex("Che.p")))
```

## Using References

The identifying functions can also return value references if specified.
Such references are needed as input for example to define Parameter Estimations.

```{r}
species("CheY", reference = "InitialConcentration")
```

These function however can not give any clues about what references are available.

```{r}
species("CheY", reference = "failure")
```

To help with this, functions implemented to get model values have similar functions that return references instead of values.

```{r}
str(getSpecies("CheY"))
str(getSpeciesReferences("CheY"))
```
