---
title: "Issues"
author: "Jonas Förster & Jürgen Pahle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Issues}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
set.seed(1)
```

## Initial Setup

```{r setup}
library(tidyverse)
library(CoRC)
```

## Checking validity of quantity unit doesn't work

```{r}
CUnit <- CoRC:::CUnit
```

The CUnit class can be used for testing if user specified input is valid

```{r}
ms <- CUnit("ms")
s <- CUnit("ms * 1e3")
mmol <- CUnit("mmol")
mol <- CUnit("mmol * 1e3")

ms$buildExpression()
s$buildExpression()
mmol$buildExpression()
mol$buildExpression()
```

CUnit$getExpression("<type>") shows what Copasi makes of the input

```{r}
  ms$getExpression()
  s$getExpression()
  mmol$getExpression()
  mol$getExpression()
```

CUnit$isUnitType("<type>") will return useful information on validity for all but "quantity"

```{r}
s$isUnitType("time")
s$isUnitType("quantity")

mol$isUnitType("time")
mol$isUnitType("quantity")
mol$isUnitType("volume")
mol$isUnitType("area")
mol$isUnitType("length")
```

## ObjectDisplayName seems more apropriate than CN

```{r}
loadExamples(1)
species("X")
species("X", "InitialConcentration")
```

When a species is identified I have to return the CommonName as reference.
In this use case, returning the ObjectDisplayNames seems more friendly to the user.
Especially because this is the way it is presented in the GUI too.

```{r}
CoRC:::cn_to_object(species("X"), datamodel = getCurrentModel())[[1]]$getObjectDisplayName()
CoRC:::cn_to_object(species("X", "InitialConcentration"), datamodel = getCurrentModel())[[1]]$getObjectDisplayName()
```

They should be able to accomplish the same job in uniquely identifying a reference but there are no methods to resolve from DisplayName back to an Object.

## Reliably setting values

I don't understand how to consistently keep the math up to date with the values set in the model.

For reaction parameter, my methods are currently broken.
I can set the value, but after running a task the value will be restored.


```{r}
loadExamples(1)
ts1 <- runTimeCourse()
getParameters("(R1).k1")$value
setParameters("(R1).k1", value = 99)
getParameters("(R1).k1")$value
ts2 <- runTimeCourse()
getParameters("(R1).k1")$value
identical(ts1, ts2)
unloadModel()
```

In this example, the following is not the right approach?:
```{r, eval=FALSE}
model()$updateInitialValues(changedObjects)
```

## Saving .cps loses experimental data

I commonly want to save a .cps to a tmp folder but when I do that experimental data paths will fail.

```{r}
datamodel <- loadSBML("http://www.ebi.ac.uk/biomodels-main/download?mid=BIOMD0000000010")

# define the experiments for copasi
experiment <- defineExperiments(
  filename = "issue_expdata",
  data = suppressMessages(read_tsv("data/MAPKdata.txt")),
  type = c("time", "dependent", "dependent"),
  mapping = c(NA, species(c("Mos-P", "Erk2-P"), "Concentration"))
)

addExperiments(experiment)

fitproblem <- as(datamodel$getTask("Parameter Estimation")$getProblem(), "_p_CFitProblem")
filenames <- fitproblem$getExperimentSet()$getFileNames()
```

The filenames are relative to the model and in the correct folder.

```{r}
filenames
file.exists(filenames)
```

Now I will save the model into a tmp folder.

```{r}
file <- tempfile(fileext = ".cps")
saveCPS(file)
# unloadModel()

datamodel <- loadModel(file)
fitproblem <- as(datamodel$getTask("Parameter Estimation")$getProblem(), "_p_CFitProblem")
filenames <- fitproblem$getExperimentSet()$getFileNames()
```

The filenames are now supposedly in the folder where the .cps was saved but there they don't exist.

```{r}
filenames
file.exists(filenames)
```

```{r}
# unloadModel()
invisible(file.remove(file, "issue_expdata.txt"))
```

## Crashes with experimental data

Running this code multiple times crashes the R session.

```{r, message=TRUE, warning=TRUE}
loadSBML("http://www.ebi.ac.uk/biomodels-main/download?mid=BIOMD0000000010")

# define the experiments for copasi
experiment <- defineExperiments(
  filename = "issue_expdata",
  data = suppressMessages(read_tsv("data/MAPKdata.txt")),
  type = c("time", "dependent", "dependent"),
  mapping = c(NA, species(c("Mos-P", "Erk2-P"), "Concentration"))
)

addExperiments(experiment)

unloadModel()
```
```{error}
Call("R_swig_delete_CExperiment", self, PACKAGE = "COPASI")
 *** caught segfault ***
address 0x0, cause 'memory not mapped'
```

My approach for adding Experiments is:

```{r, eval=FALSE}
exp <- CExperiment()
experiment_set$addExperiment(exp)
```

some time later the experiment might get removed by calling either of:

```{r, eval=FALSE}
CRootContainer_removeDatamodel(datamodel)
```
```{r, eval=FALSE}
experiment_set$removeExperiment(0)
```

Maybe R crashes some time later because SWIG does try to delete exp too?
