---
title: "Issues"
author: "Jonas Förster & Jürgen Pahle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Issues}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
set.seed(1)
```

## Initial Setup

```{r setup}
library(tidyverse)
library(CoRC)
```

## ObjectDisplayName seems more apropriate than CN

```{r, results="hide"}
loadExamples(1)
```

```{r}
species("X")
species("X", "InitialConcentration")
```

When a species is identified I have to return the CommonName as reference.
In this use case, returning the ObjectDisplayNames seems more friendly to the user.
Especially because this is the way it is presented in the GUI too.

```{r}
cmetab_x <- CoRC:::cn_to_object(species("X"), datamodel = getCurrentModel())[[1]]
cmetab_ref_x_0 <- CoRC:::cn_to_object(species("X", "InitialConcentration"), datamodel = getCurrentModel())[[1]]

cmetab_x$getObjectDisplayName()
cmetab_ref_x_0$getObjectDisplayName()
```

They should be able to accomplish the same job in uniquely identifying a reference but there are no methods to resolve from DisplayName back to an Object.

## Reliably setting values

I don't understand how to consistently keep the math up to date with the values set in the model.

For reaction parameter, my methods are currently broken.
I can set the value, but after running a task the value will be restored.

```{r}
loadExamples(1)
ts1 <- runTimeCourse()
getParameters("(R1).k1")$value
setParameters("(R1).k1", value = 99)
getParameters("(R1).k1")$value
ts2 <- runTimeCourse()
getParameters("(R1).k1")$value
identical(ts1, ts2)
unloadModel()
```

In this example, the approach used for setting species values is probably not the right one:

```{r, eval=FALSE}
model()$updateInitialValues(changedObjects)
```

## Checking validity of quantity unit doesn't work

```{r}
CUnit <- CoRC:::CUnit
```

The CUnit class can be used for testing if user specified input is valid

```{r, results="hide"}
ms <- CUnit("ms")
s <- CUnit("ms * 1e3")
mmol <- CUnit("mmol")
mol <- CUnit("mmol * 1e3")
mMol <- CUnit("mMol")

ms$buildExpression()
s$buildExpression()
mmol$buildExpression()
mol$buildExpression()
mMol$buildExpression()
```

CUnit$getExpression() shows what Copasi makes of the input.

```{r}
ms$getExpression()
s$getExpression()
mmol$getExpression()
mol$getExpression()
mMol$getExpression()
```

CUnit$isUnitType("\<type\>") will return useful information on validity for all but "quantity".

```{r}
ms$isUnitType("time")

s$isUnitType("time")
s$isUnitType("quantity")

mmol$isUnitType("quantity")

mol$isUnitType("quantity")
mol$isUnitType("time")
mol$isUnitType("volume")
mol$isUnitType("area")
mol$isUnitType("length")

mMol$isUnitType("quantity")
```

## Function naming is lacking a cohesive concept 

Quantity accessors are named:

```{r, eval=FALSE}
getSpecies()
setSpecies()
getGlobalQuantities()
setGlobalQuantities()
setParameters()
```

Task related functions are named:

```{r, eval=FALSE}
runTimeCourse()
runSteadyState()
runParamEst()
setTimeCourseSettings()
setSteadyStateSettings()
setParamEstSettings()
```

There are further functions which with awkward names:

```{r, eval=FALSE}
# Adds FitItem to parameter estimation task
# No relation to setParameters(), which handles reaction parameters
addParameter()
```

Naming so far has been based on the desire to stay as close to the GUI as possible.
It's probalby time to define a cohesive concept.

## Saving .cps loses experimental data

I commonly want to save a .cps to a tmp folder but when I do that experimental data paths will fail.

```{r}
# Load example which has experimental data.
datamodel <- loadExamples(4)[[1]]

fitproblem <- as(datamodel$getTask("Parameter Estimation")$getProblem(), "_p_CFitProblem")
filenames <- fitproblem$getExperimentSet()$getFileNames()
```

The filenames are in the correct folder.

```{r}
filenames
file.exists(filenames)
```

Now I will save the model into a tmp folder.

```{r}
file <- tempfile(fileext = ".cps")
saveCPS(file)
unloadModel()

datamodel <- loadModel(file)
fitproblem <- as(datamodel$getTask("Parameter Estimation")$getProblem(), "_p_CFitProblem")
filenames <- fitproblem$getExperimentSet()$getFileNames()
```

The filenames are now supposedly in the folder where the .cps was saved but there they don't exist.

```{r}
filenames
file.exists(filenames)
```

```{r}
unloadModel()
invisible(file.remove(file))
```
