---
title: "Task Management"
author: "Jonas Förster & Jürgen Pahle"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing Tasks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.asp = 10/16,
  out.width = "100%",
  error = TRUE,
  warning = FALSE,
  message = FALSE
)
set.seed(1)
```

# Initial Setup

```{r setup}
library(CoRC)
loadExamples(2)
```

# Task functions

CoRC has implemented a variety of COPASI functions.
Tasks generally have an `run` execution function (e.g. `runTimeCourse`), a settings `set` function (e.g. `setTimeCourseSettings`) and a settings `get` function (e.g. `getTimeCourseSettings`).

**Get task settings**:

Get functions return a list of settings in the same format as is expected as input to `set` functions.

```{r}
str(getTimeCourseSettings())
```

**Set task settings**:

```{r}
setTimeCourseSettings(duration = 100)
getTimeCourseSettings()$duration
```

**Run task**:

Run functions return a list of results.
Included in this list is an entry `settings` with contains all tasks settings.

```{r}
str(runTimeCourse())
```

# Task settings

Task settings can be set in two ways.
The `set` familiy of functions is the preferred way to set task settings.

```{r}
getTimeCourseSettings()$duration
setTimeCourseSettings(duration = 10)
getTimeCourseSettings()$duration
```

The `run` familiy of functions will generally take the same arguments as the `set` familiy of functions but will only set them temporarily.

```{r}
getTimeCourseSettings()$duration
runTimeCourse(duration = 1)$settings$duration
getTimeCourseSettings()$duration
```

The `get` familiy of functions will return a list that can be used as a template for calling functions that expect settings as input.
A similar list is also given in every output of a `run` function.

```{r}
method_settings <- getTimeCourseSettings()$method
str(method_settings)

setTimeCourseSettings(method = method_settings)
```

# Task execution

Executing a task will return a list containing all relevant task results.

```{r}
str(runTimeCourse())

str(runSteadyState())
```
