---
title: "Managing Tasks"
author: "Jonas Förster & Jürgen Pahle"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing Tasks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  out.width = "100%",
  error = TRUE,
  warning = FALSE,
  message = FALSE
)
set.seed(1)
```

## Initial Setup

```{r setup}
library(tidyverse)
library(CoRC)
loadExamples(2)
```

## Task functions

# Task functions

CoRC has implemented a variety of copasi functions.
Tasks generally have an execution function (e.g. \code{runTimeCourse}), a settings setter function (e.g. \code{setTimeCourseSettings}) and a settings getter function (e.g. \code{getTimeCourseSettings}).

```{r}
tc_settings <- getTimeCourseSettings()
tc_settings$duration

setTimeCourseSettings(duration = 100)
getTimeCourseSettings()$duration

runTimeCourse()
```

# Task settings

Task settings can be set in two ways.
The set familiy of functions is the preferred way to set task settings.
The run familiy of functions will generally take the same arguments as the set familiy of functions but will only set them temporarily.
The get familiy of functions will return a list that can be used as a template for calling functions that expect settings as input.
A similar list is also given in every output of a run function.

```{r}
getTimeCourseSettings()$duration
setTimeCourseSettings(duration = 10)
runTimeCourse()$settings$duration
runTimeCourse(duration = 1)$settings$duration
getTimeCourseSettings()$duration
```

# Task execution

Executing a task will return a list with all relevant outputs.

```{r}
str(runTimeCourse())
str(runSteadyState())
```

# Further remarks

Settings from get functions should serve as input for set functions

```{r}
settings <- getTimeCourseSettings()
do.call(setTimeCourseSettings, settings)
```

